---
title: "Tidying_Labor-Werte_from_PDF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidying_Labor-Werte_from_PDF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mydailyhelpers)
```

Required libraries:

```{r eval=FALSE}

library(tidyverse)
library(pdftools)

```



Sometimes when carrying out a study, there is a need to process the laboratory information of many patients over  the period of the study. **After getting the approval of the ethics committee and the appropiate patient consent** download the required lab results, and print as a pdf.

```{r eval=FALSE}

# Here enter the path where you stored the pdfs

pdf_paths <- dir("./patienten_suche/labor_werte/labor_pdfs",
                 full.names = TRUE,
                 pattern = "pdf")
```


the first function: mydailyhelpers::prep_labor_pdf_pages() takes the following functions:

1. path to pdf.
2. Regex-expression representing the relevant words representing the columns 
3. Regex-expression representing the empty fields. Will be returned as NA
4. Regex-expression containing the labor parameters wanted. E. g. creatinin, glukose. This is case sensitive.
5. Regex-expression representing the upper limit to include. This is usually the name of the left most column from 2.
6. Regex-expression representing a term, as the lower limit to include (for pages containing both vaild results and comments)


```{r eval=FALSE}
reg_filt <- regex(
  "^Auftragsdatum$|bereich|(?<!\\*)(\\d{2}\\.){2}\\d\\d")

empty_filt <- regex("\\.{6,14}")

y_up <- regex("^Auftragsdatum$")
st_term <- regex("Kommentar")
wanted_lab_filter <- regex("Glukose|Hb\\sA1c|Hb\\sA1C|Creatinin|Quick|PTT")
```


Running the function on a single PDF file returns a list of labor parameters with date_time per page of the original PDF

```{r eval=FALSE}
pdf_list <- pdftools::pdf_data(pdf_paths[1])

tab_list <- 
    pdf_list %>% 
    future_map(prep_labor_pdf_pages, 
        reg_filt,
        empty_filt,
        wanted_lab_filter,
        y_up, 
        st_term)

```

The resulting list, shold be combined across all pages using the second function in the package

```{r eval=FALSE}
combined_list <- tab_list %>% combine_lab_pages()

```


Alternatively all PDFS in a directory could be read and evaluated. Iteration using the map() function from the purrr package can be used. This produces a nested list.

```{r eval=FALSE}
big_out <- pdf_paths %>% 
  set_names(nm = str_extract(., "\\d{8}")) %>% # pass personal id nummber as names for the list elements
  
  map(function(path){
  pdf_list <- pdf_data(path)
  tab_list <- 
    pdf_list %>% 
    future_map(prep_labor_pdf_pages, 
        reg_filt,
        empty_filt,
        wanted_lab_filter,
        y_up, 
        st_term)
  combined_list <- tab_list %>% combine_lab_pages()
  }
  )
```


the following code combines all results from all patients into one tibble, to allow wrangling and filtering.

first, bind_rows() is mapped to the inner list, then reapplied to the outer list to produce the tidy table

```{r eval=FALSE}
big_out %>% map_depth(1, bind_rows, .id = "labor_param") %>% 
  bind_rows(.id = "piz")
```






